---
title: Detector Dog CBARQ with Phase 1 and Phase 2 Data
subtitle: "V 1.0 Both Data Sets and No Chasing Questions"
author: Liz Hare
date: "`r Sys.Date()`"
email: "lizhare@gmail.com"
output:
    html_document:
        toc: true
        toc_depth: 5
        number_sections: true
---

``` {r, setup, include=FALSE, echo=FALSE}
### load packages
library(tidyverse)
library(lubridate)
library(janitor)
library(stringr)
library(knitr)
library(kableExtra)
library(gtsummary)
library(writexl)
library(psych)
library(correlation)
library(gluedown) ## inline print lists from R
library(openxlsx)
library(writexl)


### get files

rawPhase1 <- read.xlsx("data/Detection-dog-survey.xlsx", sheet = 1)
rawPhase2 <- read.xlsx("data/DD Survey 20220714-20221104.xlsx", sheet = 1)
descItems <- read.csv("data/cbarqItemDescriptions.csv")

```

# Data and Code

The R code for the analysis of the combined Phase 1 and Phase 2 data are in a 
[GitHub repository](https://github.com/LizHareDogs/detectorCBARQ).
This repository is private for now; send me a request to access it with your GitHub
user ID at <lizhare@gmail.com>. The goal is open the code for peer review
and the community.  

Two spreadsheets received from James are imported:

- Phase 1: `Detection-dog-survey.xlsx`
- Phase 2: `DD Survey 20220714-20221104.xlsx`

``` {r, cleanDemog, results="asis", echo=FALSE}
source("demogPhase1.R")
source("demogPhase2.R")

d1 <- bind_rows(rawPhase1, rawPhase2, .id = "phase")

### keep only the last date for each dog to remove duplicates
d1 <- d1 %>%
    group_by(Name, Email) %>%
    slice_max(surveyDate, n = 1) %>%
    ungroup()

### check again for duplicates
duplicatedDogs <- nrow(d1) - nrow(unique(d1[ ,c("Name", "Email")]))
dupdogsDF <- get_dupes(d1, c(Name, Email))
write.csv(dupdogsDF, file="dupdogs.csv", quote=FALSE, row.names=FALSE)

### two dogs have records from the same day
### count NA per row for each pair
##rowSums(is.na(dupDogsDF))
## all.equal(dupDogsDF[1,], k[2, ])
## all.equal(dupDogsDF[3, ], k[4, ])

### delete by assigned ID number for records with lots of NAs
### Dog.ID is per record, so extra records can be delted by their ID numbers
delRecs <- c(162150, 155420)
d1 <- d1[!d1$Dog.ID %in% delRecs, ]
save(d1, file = "d1.Rda")

```


# Demographic Description

The data analyzed here include
`r nrow(d1)` dogs with dogs collected in Phase 1
(`r nrow(d1[d1$phase == 1,])` dogs) and Phase 2 
(`r nrow(d1[d1$phase == 2, ])` dogs). In phase 1, there was a problem collecting 
data about whether dogs were neutered. Also in phase 1,
three questions about chasing were excluded. Because they are 
missing fro most dogs in this data set, they are excluded from the 
whole analysis (previous analysis of Phase 2 shows the three chasing items
separate into a factor).  

After merging the data from Phase 1 and Phase 2, we found that 58 dogs 
had repeated records. 56 were removed by selecting the record with the
latest date, because they will have better data on neutering. Two 
repeated records were completed on the same day, so the ones with fewer
missing CBARQ items were retained.  

## Sex and Neuter Status

``` {r, sexNeutered, results="asis", echo=FALSE}

sexNeuteredTab <- tbl_summary(d1, by = neuteredF,
                             include = sexF,
                             type = everything() ~ "categorical",
                             statistic = everything() ~ "{n} ({p}%)",
                             missing = "ifany") %>%
add_overall()

modify_caption(sexNeuteredTab, "Overall Sex and Neuter Status (Caution: lots of missing neuter status data)")

### break out by phase

sexNeuteredPhaseTab <- tbl_strata(d1,
                                  strata = phase,
                                  .tbl_fun = ~ .x %>%
                                      tbl_summary(by = sexF,
                                                  include = "neuteredF",
                                                  statistic = everything() ~ "{n} ({p}%)",
                                                  missing = "ifany") %>%
modify_caption("Sex and Neuter Status by Phase(Caution lots of missing neuter status data in Phase 1_"))
sexNeuteredPhaseTab
                                                 
```

## Breed and Breed Group

The questionnaire provided a multiple-choice question asking about the 
dog's breed. We assigned breeds to 
[AKC breed groups](https://www.akc.org/public-education/resources/general-tips-information/dog-breeds-sorted-groups/) to potentially avoid having a large number of categories with small numbers
of dogs. If a breed was not listed with AKC, breeds, groups were chosen according to the breed's 
purpose and history.

Overall, there were `r nrow(d1[!is.na(d1$breedGroup), ])` purebred dogs.
There were 
`r nrow(d1[d1$Breed == "Cross-breed", ])` dogs listed as 
"cross-breed" (*were there instructions to use this specifically
for known crosses like labradoodles?*) and 
`r nrow(d1[d1$Breed == "Mixed-breed", ])` dogslisted as "mixed-breed."  


Cross-breed
Mixed-breed

### Breed

``` {r, breed, results="asis", echo=FALSE}

breedTab <- tbl_summary(d1,
                        include = Breed,
                        statistic = all_categorical() ~ "{n} ({p}%)")
breedTab

### for Excel:
breedTabOut <- modify_header(breedTab,
                             label = "Breed",
                             stat_0 = "N (1,117) (Percent)") %>%
    gtsummary::as_tibble()
write_xlsx(breedTabOut, "TableS4.xlsx")

``` 

### Breed Group

``` {r, breedGroup, results="asis", echo=FALSE}
breedGroupTab <- tbl_summary(d1,
                             include = breedGroup,
                             statistic = all_categorical() ~ "{n} ({p}%)",
                             digits = list(all_categorical() ~ c(0,1))) %>%
modify_header(
    label = "Breed Group",
    stat_0 = "N (Percent)")

breedGroupTab

breedGroupTib <- gtsummary::as_tibble(breedGroupTab)
breedGroupTib <- breedGroupTib[2:nrow(breedGroupTib), ]
write_xlsx(breedGroupTib, "Table1.xlsx")


```

## Country

``` {r, country, results="asis", echo=FALSE}
countryTab <- tbl_summary(d1,
                          include = Country,
                          statistic = all_categorical() ~ "{n} ({p}%)")
countryTab

```

## Age, Training, and Retirement

``` {r, ageTrainingRet, restults="asis", echo=FALSE}
s1 <- psych::describe(d1$ageY)
s1 <- round(s1, 2)
s1K <- kable(s1, caption="Univariate Statistics for Age in Years")
kable_styling(s1K)

ageWorkStatusTab <- tbl_summary(d1, by=workingStatusF,
                                include=ageY,
                                type = everything() ~ "continuous",
                                statistic = all_continuous() ~ "{mean} ({sd}) ({min} - {max})",
                                digits = list(everything() ~ c(2, 2, 2, 2)))
add_overall(ageWorkStatusTab)

```
 
### Dual Purpose


There are `r nrow(d1[d1$duallyTrainedF == "yes", ])` dual purpose dogs
(`r round(nrow(d1[d1$duallyTrainedF == "yes", ])/nrow(d1), digits=2)*100`%).

#### What breeds are dually trained?

From some of the breeds that show up in this table, I suspect these handlers
aren't defining "dual purpose" the same way PVWDC does (Golden Retriever, 
Cocker Spaniel...).  


``` {r, dualTrain, results="asis", echo=FALSE}
d1Dual <- d1[d1$duallyTrainedF == "yes",]
breedDualTab <- tbl_summary(d1Dual, include=breedF,
                            statistic = all_categorical() ~ "{n}, ({p}%)",
                            digits = list(breedF ~ c(0, 1)))
breedDualTab
```

## Odors Detected

``` {r, odorVars, results="asis", echo=FALSE}


odorTab <- tbl_summary(d1, 
                       include=odoraccelerant:odorother,
                       type=everything() ~ "dichotomous",
                       statistic = all_categorical() ~ "{n} ({p}%)",
                       digits = list(all_categorical() ~ c(0, 1)),
                       label = list(odoraccelerant = "Accelerant",
                                    odorcontraband = "Contraband",
                                    odorendangeredspecies = "Endangered Species",
                                    odorexplosives = "Explosives",
                                    odorhumanremains = "Human Remains",
                                    odorinvasivespecies = "Invasive Species",
                                    odorlivehumanairscent = "Live Human Air Scent",
                                    odorlivehumantracking = "Live Human Tracking",
                                    odormedicaldetection = "Medical Detection",
                                    odornarcotics = "Narcotics",
                                    odorpests = "Pests",
                                    odorother = "Other")) %>%
    modify_header(
        label = "Type of Trained Odor",
        stat_0 = "N (Percent)")       
odorTab

odorTib <- gtsummary::as_tibble(odorTab)
write_xlsx(odorTib, "Table2.xlsx")




```

### Odors entered under "Other"

There are too many responses to print here 
(`r nrow(d1[d1$odorother == "yes" & !is.na(d1$odordescription), ])`). 
Lots of them are nosework-related. Let me know if you want the list.

## Certification

``` {r, certTab, results="asis", echo=FALSE}
certTab <- tbl_summary(d1,
                       include=certificatearda:certificateuspda,
                       type=everything() ~ "dichotomous",
                       statistic = all_categorical() ~ "{n} ({p}%)",
                       digits = list(everything() ~ c(0, 1)),
                       label = list(certificatearda = "American Rescue Dog Association",
                                     certificatefema = "Federal Emergency Management Agency",
                                     certificatenasr = "National Association for Search and Rescue",
                                     certificatenesdca = "National Entomology Scent Detection Canine Association",
                                     certificatenndda = "National Narcotic Detector Dog Association",
                                     certificatentpda = "National Tactical Police Dog Association",
                                     certificatenapda = "North American Police Dog Association",
                                     certificatesrdus = "Search and Rescue Dogs of the United States",
                                     certificateuspda = "United States Police Dog Association",
                                     certificateother = "Other")) %>%
                           modify_header(
                               label = "Certifying Organization",
                               stat_0 = "N (Percent)")


                       certTab

                       certTib <- gtsummary::as_tibble(certTab)
                       write_xlsx(certTib, "Table3.xlsx")

```

### Other Certifying organizations

There are too many answers to print here
(`r nrow(d1[d1$certificateother == "yes" & !is.na(d1$certificatedescription), ])`).  

## Alert Type

``` {r, alertTab, results="asis", echo=FALSE}
d1$alerttypeF <- str_to_title(d1$alerttypeF)
alertTab <- tbl_summary(d1, include=alerttypeF,
                        type = everything() ~ "categorical",
                        statistic = all_categorical() ~ "{n}, ({p}%)",
                        digits = list(all_categorical() ~ c(0,1))) %>%
    modify_header(
        label = "Alert Type",
        stat_0 = "N (Percent)")

alertTab

alertTib <- gtsummary::as_tibble(alertTab)
alertTib <- alertTib[2:nrow(alertTib),]
write_xlsx(alertTib, "Table4.xlsx")

```

## Reward System

``` {r, rewardTab, results="asis", echo=FALSE}
rewardTab <- tbl_summary(d1, include=rewardsystemF,
                         type = everything() ~ "categorical",
                         statistic = all_categorical() ~ "{n}, {p}%",
                         digits = list(all_categorical() ~ c(0,1)))
rewardTab
```

## Handler Experience

Experience is measured by the number of dogs handled before the current one.

``` {r, dogsHandledTab, results="asis", echo=FALSE}
dogsHandledTab <- tbl_summary(d1, include=dogshandledbeforeF,
                              type = everything () ~ "categorical",
                              statistic = all_categorical() ~ "{n}, {p}%",
                              digits = list(all_categorical() ~ c(0,1)))
dogsHandledTab
```

## Dogs' Home Environment

``` {r, dogsHome, results="asis", echo=FALSE}
dogsHomeTab <- tbl_summary(d1, include=dogshomelocationF,
                           type = everything () ~ "categorical",
                           statistic = all_categorical() ~ "{n}, {p}%",
                           digits = list(all_categorical() ~ c(0,1)))

dogsHomeTab
```
## Owner Rating of Scent Work Ability and General Behavior and Temperament

### Scent Work

``` {r, scenntWorkDesc, results="asis", echo=FALSE}

scentDesc <- psych::describe(d1$scentrating)
scentDesc2 <- round(scentDesc, 2)
scentDescK <- kable(scentDesc2,
                    caption = "Owners' Rating of Dogs' Scent Detection Work Ability")
kable_styling(scentDescK)
```

### General Behavior and Temperament

``` {r, behaviorRate, results="asis", echo=FALSE}
behavDesc <- psych::describe(d1$behaviorrating)
behavDesc <- round(behavDesc, 2)
behavDescK <- kable(behavDesc,
                    caption = "Owners' Rating of Dogs' General Behavior and Temperament")
kable_styling(behavDescK)
```

### Are Scent Work and Behavior Assessment Correlated?

``` {r, assessmentCorrelation, results="asis", echo=FALSE}
assessmentCorrelation <- cor(d1$scentrating, d1$behaviorrating,
                             use = "pairwise.complete.obs")

```
The correlation between scenting ability and general behavior is
`r round(assessmentCorrelation, 2)`.

## DNA Sample Availability

There were `r nrow(d1[d1$sampleprovidedF == "yes", ])` owners willing to provide DNA samples 
(`r round(nrow(d1[d1$sampleprovidedF == "yes", ])/nrow(d1)*100)`%).


# CBARQ Items' Descriptive Statistics and Plots

``` {r, cleanNames, results="asis", echo=FALSE}
ddItems <- select(d1, c(TRAIN01:AGG20, FEAR21:SEPR34, EXCITE35:EXCITE38,
                         ATT39:ATT42, PLAY43:MISC71))
### are they all numeric?
### table(sapply(ddItems, class))
### do they have any weird outliers
### sapply(ddItems, table, useNA="ifany")
trainItems <- select(d1, TRAIN01:TRAIN08)
### class(d1$othersituationsagg)
### above variable character so not plotted-- all "othersituations* vars
aggItems <- select(d1, AGG09:AGG20)
fearItems <- select(d1, FEAR21:FEAR31)
sepItems <- select(d1, SEPR32:SEPR34)
excitItems <- select(d1, EXCITE35:EXCITE38)
attItems <- select(d1, ATT39:ATT42)
playItems <- select(d1, PLAY43:PLAY46)
impItems <- select(d1, IMP47:IMP49, IMP50)
miscItems <- select(d1, MISC51:MISC71)

```
 



``` {r, describeItems, results="asis", echo=FALSE}


save(ddItems, file="data/ddCBARQnumeric.Rda")

### summary stats
### count missing

countMissing <- sapply(ddItems, function(x) sum(is.na(x)))
means <- round(sapply(ddItems, mean, na.rm=TRUE), 2)
sds <- round(sapply(ddItems, sd, na.rm=TRUE), 2)
mins <- sapply(ddItems, min, na.rm=TRUE)
maxes <- sapply(ddItems, max, na.rm=TRUE)
skews <- round(sapply(ddItems, skew, na.rm=TRUE), 2)
kurts <- round(sapply(ddItems, kurtosi, na.rm=TRUE), 2)
### shapiro-wilk test for normality
sws <- round(sapply(ddItems, function(x) shapiro.test(x)$statistic), 2)
swp <- format(sapply(ddItems, function(y) shapiro.test(y)$p.value), scientific = TRUE)

bigDesc <- data.frame(countMissing, means, sds, mins, maxes, skews, kurts, sws, swp)
colnames(bigDesc) <- c("NMissing", "Mean", "SD", "Min", "Max", "Skewness", "Kurtosis",
                       "Shapiro-Wilk Score", "Shapiro-Wilk p")

bigDescK <- kable(bigDesc,
      caption="Descriptive Statistics for Detector-C-BARQ Items")

kable_styling(bigDescK)

### make output Excel table
bigDesc$itemNames <- rownames(bigDesc)
bigDesc <- bigDesc[ , c(ncol(bigDesc), 2:ncol(bigDesc) - 1)]
colnames(bigDesc) <- c("Item", "Number of Missing Items", "Mean", "Standard Deviation",
                       "Minimum", "Maximum", "Skewness", "Kurtosis", "Shapiro-Wilk Statistic",
                       "Shapiro-Wilk P-value")
write_xlsx(bigDesc, "TableS5.xlsx")


largeSkew <- bigDesc[bigDesc$skews < -2 & bigDesc$skews > 2, ]

largeSkewItems <- rownames(bigDesc[bigDesc$Skewness < -2 | bigDesc$Skewness  > 2, ])
largeKurtItems <- rownames(bigDesc[bigDesc$Kurtosis > 7, ])
### all items you'd remove
removeSK <- unique(c(largeSkewItems, largeKurtItems))


```

The number of dogs in this analysis is `r nrow(ddItems)`. To use mean or mode imputation, < 5% of the data should be missing. 
Items should have no more than `r nrow(ddItems)*0.05` missing if these kinds of imputations will be used. 

``` {r, Missing5Percent, results="asis", echo=FALSE}

highMissing <- bigDesc[bigDesc$NMissing > nrow(d1)*0.05, ]
highMissingK <- kable(highMissing, digits=2,
                      caption = "Items with > 5 percent missing")
kable_styling(highMissingK)

```

There are `r nrow(highMissing)` items with > 5\% missing.  


According to Watkins (2021), items should be removed if they have skewness > 2.0 or 
kurtosis > 7.0. These items would violate the assumptions of the Pearson correlation coefficients 
that factor analysis is based on. (Do we need to consider the items as ordinal discrete values?)  

There are `r length(largeSkewItems)` items with skewness > 2.0 including 
`r largeSkewItems`.  

There are `r length(largeKurtItems)` items with kurtosis > 7.0 including
`r largeKurtItems`

For the current data set, `r length(removeSK)`
items would be removed for the factor analysis. 

### Count Plots

#### Trainability


``` {r, plotTrainFreq, results="asis", echo=FALSE, fig.width=12, fig.height=24}
### plot



plotFreq <- function(x) {
    longData <- pivot_longer(x, cols = everything(),
                             names_to = "item", values_to = "score")
    ggplot(longData, aes(x = score)) +
                     geom_bar(na.rm=TRUE) +
                     facet_wrap(vars(item), ncol=4) +
                     labs(x = "Score", y="Count",
                          title="Counts")
    }
plotFreq(trainItems) 
```

#### Aggression

``` {r, plotAggFreq, results="asis", echo=FALSE}
plotFreq(aggItems)
```

#### Fear

``` {r, plotFearFreq, results="asis", echo=FALSE}
plotFreq(fearItems)
```


#### Separation

``` {r, plotSepFreq, results="asis", echo=FALSE}
plotFreq(sepItems)
```


#### Excitability

``` {r, plotExcitFreq, results="asis", echo=FALSE}
plotFreq(excitItems)
```

#### Attention-Seeking

``` {r, plotAttFreq, results="asis", echo=FALSE}
plotFreq(attItems)
```



#### Play

``` {r, plotPlayFreq, results="asis", echo=FALSE}
plotFreq(playItems)
```


#### Impulsivity

``` {r, plotFrustFreq, results="asis", echo=FALSE}
plotFreq(impItems)
```


#### Miscellaneous

``` {r, plotMiscFreq, results="asis", echo=FALSE}
plotFreq(miscItems)
```


# Conclusions for Factor Analysis

Data will ba analyzed with the items as **ordered factors** rather than numeric
variables for the following reasons:

- Many items have a large number of values of 0  
- The p-value for the Shapiro test is always near 0, indicating that 
the data do not come from a normal distribution.  
- Some items have high skewness and kurtosis. The kurtosis is particularly high 
for some of the items.

``` {r, convertToOrderedAndSave, results="asis", echo=FALSE, warnings=FALSE}
### calculating correlation matrixir requires numeric fields even if
###                                                        you are going to analyze them as ordered
save(ddItems, file="data/ddCBARQnumeric.Rda")

### use lapply because sapply ended up with character vars
### input for imputation needs it in a data frame
ddItemsOrdered <- data.frame(lapply(ddItems, as.ordered))

save(ddItemsOrdered, file="data/ddCBARQordered.Rda")



```
