## {{nfact}} Factors

### {{nfact}} Factors Model Fit

``` {r, modelFit{{nfact}}, eval=TRUE, echo=FALSE}

fanal{{nfact}} <- fa(imputedNumericDF,
      nfactors = {{nfact}},
      n.obs = nrow(imputedNumericDF),
      rotate = "oblimin",
      fm = "pa",
      smooth = TRUE,
      cor = "poly",
      correct = 0.01)

save(fanal{{nfact}}, file = paste0("data/fanal", {{nfact}}, ".Rda"))

```

Although the chi-square test of goodness of fit is sensitive to departures from normality like
the C-BARQ items, Hopper et al (2008) recommend always reporting it.  

- chi-square: `r fanal{{nfact}}$STATISTIC`  
- degrees of freedom: `r fanal{{nfact}}$dof`  
- P-value for chi-square = `r fanal{{nfact}}$PVAL`  

Tucker-Lewis Index of Factoring Reliability/Non-Norm Fit Index:
`r fanal{{nfact}}$TLI`.
Should be > 0.9; need reference)


### {{nfact}} Factor Model Communalities

``` {r, commun{{nfact}}, eval=TRUE, echo=FALSE}
comTab <- kable(data.frame(fanal{{nfact}}$communality), digits = 2)
kable_styling(comTab, full_width = FALSE)

```

#### How many communalities < 0.40?

There are `r length(fanal{{nfact}}[fanal{{nfact}}$communality < 0.40])` 
items with communality < 0.40.


``` {r, lowCommTab{{nfact}}, eval=TRUE, echo=FALSE}
lc{{nfact}} <- data.frame(fanal{{nfact}}$communality)
lc{{nfact}}$item <- colnames(ddItems)
lc{{nfact}} <- lc{{nfact}}[lc{{nfact}}[1] < 0.40, ]
lcTab{{nfact}} <- kable(lc{{nfact}}, digits = 2)
kable_styling(lcTab{{nfact}}, full_width = FALSE)
datapath <- "data/"
save(lc{{nfact}}, file=paste0(datapath, "lc", {{nfact}}, ".Rda"))

```

#### How many communalities < 0.50?

There are `r length(fanal{{nfact}}[fanal{{nfact}}$communality < 0.50])` 
items with communality < 0.50.


``` {r, lowCommTabF{{nfact}}, eval=TRUE, echo=FALSE}
lcf{{nfact}} <- lc{{nfact}}[lc{{nfact}}[1] < 0.50,]
save(lcf{{nfact}}, file=paste0(datapath, "lc", {{nfact}}, ".Rda"))

lcfTab{{nfact}} <- kable(lcf{{nfact}}, digits = 2)
kable_styling(lcfTab{{nfact}}, full_width = FALSE)

```




### {{nfact}} Factor Model Loadings


### All loadings

``` {r, finalLoadings{{nfact}}, eval=TRUE, echo=FALSE}
loadMat <- matrix(fanal{{nfact}}$loadings, nrow = ncol(ddItems))
dimnames(loadMat) <- dimnames(fanal{{nfact}}$loadings)
loadingsTab <- kable(loadMat, digits = 2)
kable_styling(loadingsTab)

### calculate max loading for each item
loadDF <- data.frame(loadMat)
loadDF$maximum <- apply(loadDF, 1, max)
loadDF$largest <- colnames(loadDF)[apply(loadDF, 1, which.max)]
### sort df by factor
loadDF <- loadDF[order(loadDF$largest), ]
```

### {{nfact}} Factor Model Largest Loading Per Item

``` {r, loadwithDesc{{nfact}}, eval=TRUE, echo=FALSE}
## merge description with loadings
loadDF$itemNames <- rownames(loadDF)
m1 <- merge(loadDF, descItems, by="itemNames", all.x=TRUE, all.y=FALSE)
### sort by factor 

m1 <- m1[order(m1$largest), ]
largestLoadings <- kable(m1[ ,c("largest", "maximum", "itemNames", "itemDescriptions")],
                         digits = 2,
                         caption = "Largest Loading Per Item and Associated Factors, Sorted by Factor")

kable_styling(largestLoadings, full_width = FALSE)

```

#### Items with Loadings < 0.32

These are probably too low to include

``` {r, low{{nfact}}, eval=TRUE, echo=FALSE}
m1small{{nfact}} <- m1[m1$maximum < 0.32, ]
m1small{{nfact}}$fnum <- str_extract(m1small{{nfact}}$itemNames, "\\d\\d")
m1small{{nfact}} <- m1small{{nfact}}[order(m1small{{nfact}}$fnum), c("fnum", "itemNames", "largest", "itemDescriptions")]
m1small{{nfact}}K <- kable(m1small{{nfact}}, digits = 2)
kable_styling(m1small{{nfact}}K, full_width=FALSE)


```

### {{nfact}} Model Reliability

``` {r, reliability{{nfact}}, eval = TRUE, echo=FALSE}
keys{{nfact}} <- factor2cluster(fanal{{nfact}})
keysList{{nfact}} <- keys2list(keys{{nfact}})

scores{{nfact}} <- scoreItems(keysList{{nfact}},
                              imputedNumericDF)

alpha{{nfact}}Tab <- data.frame(t(scores{{nfact}}$alpha))
alpha{{nfact}}Tab$fnum <- str_extract(rownames(alpha{{nfact}}Tab),
                                   "\\-*\\d+\\.*\\d*")
alpha{{nfact}}Tab$fnum <- str_pad(alpha{{nfact}}Tab$fnum, width = 2, side = "left", pad = "0")
alpha{{nfact}}Tab$factor <- paste0("F", alpha{{nfact}}Tab$fnum)
alpha{{nfact}}Tab  <- alpha{{nfact}}Tab[order(alpha{{nfact}}Tab$fnum), ]

alpha{{nfact}}TabK <- kable(alpha{{nfact}}Tab, digits = 2)
kable_styling(alpha{{nfact}}TabK, full_width = FALSE)

rel{{nfact}} <- reliability(keysList{{nfact}},
                            imputedNumericDF)



rel{{nfact}}

```
