## {{nfact}} Factors

### {{nfact}} Factors Model Fit

``` {r, modelFit{{nfact}}, eval=TRUE, echo=FALSE}

fanal{{nfact}} <- fa(postImputedCor,
      nfactors = {{nfact}},,
      n.obs = nrow(ddItems),
      rotate = "oblimin",
      fm = "wls")
```

Although the chi-square test of goodness of fit is sensitive to departures from normality like
the C-BARQ items, Hopper et al (2008) recommend always reporting it.  

- chi-square: `r fanal{{nfact}}$STATISTIC`  
- degrees of freedom: `r fanal{{nfact}}$dof`  
- P-value for chi-square = `r fanal{{nfact}}$PVAL`  

Tucker-Lewis Index of Factoring Reliability/Non-Norm Fit Index:
`r fanal{{nfact}}$TLI`.
Should be > 0.9; need reference)


### {{nfact}} Factor Model Communalities

``` {r, commun{{nfact}}, eval=TRUE, echo=FALSE}
comTab <- kable(data.frame(fanal{{nfact}}$communality), digits = 2)
kable_styling(comTab, full_width = FALSE)

```

#### How many communalities < 0.40?

There are `r length(fanal{{nfact}}[fanal{{nfact}}$communality < 0.40])` 
items with communality < 0.40.


``` {r, lowCommTab{{nfact}}, eval=TRUE, echo=FALSE}
lc{{nfact}} <- data.frame(fanal{{nfact}}$communality)
lc{{nfact}}$item <- colnames(ddItems)
lc{{nfact}} <- lc{{nfact}}[lc{{nfact}}[1] < 0.40, ]
lcTab{{nfact}} <- kable(lc{{nfact}}, digits = 2)
kable_styling(lcTab{{nfact}})
datapath <- "data/"
save(lc{{nfact}}, file=paste0(datapath, "lc", {{nfact}}, ".Rda"))

```

### {{nfact}} Factor Model Loadings


### {{nfact}} Factor Model Loadings

``` {r, finalLoadings{{nfact}}, eval=TRUE, echo=FALSE}
loadMat <- matrix(fanal{{nfact}}$loadings, nrow = ncol(ddItems))
dimnames(loadMat) <- dimnames(fanal{{nfact}}$loadings)
loadingsTab <- kable(loadMat, digits = 2)
kable_styling(loadingsTab)

### calculate max loading for each item
loadDF <- data.frame(loadMat)
loadDF$maximum <- apply(loadDF, 1, max)
loadDF$largest <- colnames(loadDF)[apply(loadDF, 1, which.max)]
### sort df by factor
loadDF <- loadDF[order(loadDF$largest), ]
```

### {{nfact}} Factor Model Largest Loading Per Item

``` {r, loadwithDesc{{nfact}}, eval=TRUE, echo=FALSE}
## merge description with loadings
loadDF$itemNames <- rownames(loadDF)
m1 <- merge(loadDF, descItems, by="itemNames", all.x=TRUE, all.y=FALSE)
### sort by factor 

m1 <- m1[order(m1$largest), ]
largestLoadings <- kable(m1[ ,c("largest", "maximum", "itemNames", "itemDescriptions")],
                         digits = 2,
      caption = "Largest Loading Per Item and Associated Factors, Sorted by Factor")
kable_styling(largestLoadings, full_width = FALSE)

```
